---
title: "Lab 8"
author: "Pstat 174/274"
date: "May 25, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.show = "hold")
```

# Summary of `R` Commands 


- To estimate parameters of an AR model:
```{r eval = FALSE}
ar(data, aic = TRUE, order.max = NULL, method = c("..."))
```

- To estimate parameters of an MA or ARMA model: 
```{r eval = FALSE}
arima(data, order = c(p, 0, q), method = c("..."))
```

- To compare models using AICC: 
```{r eval = FALSE}
AICc(fittedModel)
```

- To difference a time series at lag $d$:
```{r eval = FALSE}      
diff(data,lag = d)
```

- To predict future observations given a fitted ARMA model: 
```{r eval = FALSE}      
predict(fittedModel, number of future observations to forecast)
```

# Model Identification, Estimation and Diagnostics

## Exploratory Data Analysis

1. Analyze the Dow Jones Index data by downloading `dowj.txt` from Gauchospace. Move the file into `R`’s working directory and load the data set into R using the command `scan("dowj.txt")`.
```{r echo=FALSE}
# Set working directory
#setwd("/Users/jzapata/Dropbox/Winter 17/TA 174-274/Labs")
```

```{r}
# Load data
dowj_data <- scan("dowj.txt")
```

2. Plot the time series. What do you notice?
```{r}
dowj <- ts(dowj_data)
# Plot data
ts.plot(dowj,main = "Dow Jones Index")
```

3. Make the data stationary. What procedures were used?
```{r,fig.height=4,fig.width=8}
op <- par(mfrow=c(1,2))
acf(dowj)
pacf(dowj)
par(op)
```

```{r}
dowj.diff <- diff(dowj,1)
ts.plot(dowj.diff, main = "De-trended data")
```

## Model Identification

4. Plot the ACF and PACF. What models do they suggest?
```{r,fig.height=4,fig.width=8}
op <- par(mfrow=c(1,2))
acf(dowj.diff)
pacf(dowj.diff)
par(op)
```

## Model Estimation

5. Fit the AR model suggested by the sample PACF and estimate the coefficients using Yule-Walker estimation.
```{r}
(fit.ar <- ar(dowj.diff, method="yw"))
```

6. Construct 95% confidence intervals for the estimated AR coefficients (Hint: obtain the asymptotic variance of the estimated coefficient from the fitted `ar()` object using `fittedObject$asy.var.coef`).
```{r}
# 95% CI for phi1
ar1.se <- sqrt(fit.ar$asy.var.coef)
c(fit.ar$ar - 1.96*ar1.se, fit.ar$ar + 1.96*1.96*ar1.se)
```

7. Fit different ARMA models using maximum likelihood estimation and compare the model fits using AICC (Hint: use `arima()` for estimation and `AICc()` in `library(qpcR)` for model comparison - you will need to install this package into `R` first). Which model is preferred?
```{r, message=FALSE}
library(qpcR)
# Calculate AICc for ARMA models with p and q running from 0 to 5
aiccs <- matrix(NA, nr = 6, nc = 6)
dimnames(aiccs) = list(p=0:5, q=0:5)
for(p in 0:5)
{
  for(q in 0:5)
  {
    aiccs[p+1,q+1] = AICc(arima(dowj.diff, order = c(p,0,q), method="ML"))
  }
}
aiccs
(aiccs==min(aiccs))
```

## Model Diagnostics

8. Perform diagnostics on the chosen model fit. Do the residuals appear to be white noise? Are they normally distributed?

```{r}
# Pick AR(1) and perform residual analysis:
fit = arima(dowj, order=c(1,1,0), method="ML")

# Test for independence of residuals
Box.test(residuals(fit), type="Ljung")
# Test for normality of residuals
shapiro.test(residuals(fit))
```

```{r,  fig.width=6, fig.height=4}
ts.plot(residuals(fit),main = "Fitted Residuals")
```

```{r, fig.width=8, fig.height=7}
par(mfrow=c(1,2),oma=c(0,0,2,0))
# Plot diagnostics of residuals
op <- par(mfrow=c(2,2))
# acf
acf(residuals(fit),main = "Autocorrelation")
# pacf
pacf(residuals(fit),main = "Partial Autocorrelation")
# Histogram
hist(residuals(fit),main = "Histogram")
# q-q plot
qqnorm(residuals(fit))
qqline(residuals(fit),col ="blue")
# Add overall title
title("Fitted Residuals Diagnostics", outer=TRUE)
par(op)
```

# Data Forecasting

9. Forecast the next 10 observations using your model.
```{r}
# Predict 10 future observations and plot
mypred <- predict(fit, n.ahead=10)
ts.plot(dowj, xlim=c(0,89))
points(79:88,mypred$pred)
lines(79:88,mypred$pred+1.96*mypred$se,lty=2)
lines(79:88,mypred$pred-1.96*mypred$se,lty=2)
```